Article Title

### 2013-12-06 plos.tex
minor
incorporated some of Andreas' comments.

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons in behaving animals.  It is anticipated that correlations measured in such recordings will uncover aspects of functional organization of neural circuits.  However, estimation and interpretation of large correlation matrices from finite recordings can be challenging.  Estimation can be improved by regularization: the biasing of the estimate toward a low-dimensional, less variable approximation.  The amount of improvement depends on how closely the reduced approximation captures the dominant interactions with fewest terms.  Therefore, the selection of the most efficient estimator is an empirical question dependent on the system under investigation.

We compared regularized correlation matrix estimators biased toward four respective families of low-dimensional correlation structures: independent, latent factors, sparse partial correlations, and sparse partial correlations with latent factors.  The estimators' performance was evaluated on the noise correlation matrices from spatially compact groups of 91--314 neurons (average 206) in mouse visual cortex (31 sites in 24 animals) acquired by \emph{in vivo} fast 3D random-access two-photon imaging of calcium signals.  Recordings lasting between 15 and 20 minutes were deconvolved and binned at 150 ms.  Each estimator was optimized and evaluated by nested cross-validation.  We found that both sparse partial correlations and latent factors were required for efficient estimation.  In addition to reducing estimation error, optimized estimates provided compact representations of the correlation structure consisting of a sparse network of pairwise interactions and global diffuse fluctuations.  The density of positive interactions in these networks decreased rapidly with distance and with difference in orientation preference, whereas negative interactions were less selective.  The physiological interpretation of these components remains in question and must be addressed by future experiments where synaptic connectivity, cell types, and local field potentials may be used to corroborate such interpretations.

@@

The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(1),\ldots,x(n)$ as
\begin{equation}
C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T, 
\end{equation}
where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent).  
The sample covariance matrix is unbiased, \emph{i.e.} $\E{C_{\sf 0}}=\Sigma$.
For finite sample sizes, however, $C_{\sf 0}$ is not as close to $\Sigma$ as a number of biased estimators.

@@

Finally, we consider the fourth estimator $C_{\sf sparse+lowrank}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D). This estimator has the form
\begin{equation}
C_{\sf sparse+latent} = (S - LL^\T)^{-1},
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a $d\times p$ matrix of factor loadings. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.

@@

\section*{Discussion}
TODO

\section*{Methods}
% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\subsection*{Data Acquisition}
Fast 3D two-photon imaging of calcium signals \emph{in vivo} was performed as described in \cite{Cotton:2013}
\subsection*{Covariance Estimation}
TODO
\subsection*{Simulation}
TODO
\subsection*{Cross-validation}
TODO

\subsection*{Evaluation}
\section*{Acknowledgments}
Genevera Allen

>>>>

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons in behaving animals.  It is anticipated that correlations measured in such recordings will uncover important aspects of the functional organization of neural circuits.  However, estimation and interpretation of large correlation matrices from finite recordings are challenging.  Estimation can be improved by regularization: the biasing of the estimate towards a low-dimensional, less variable approximation.  The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies.  Therefore, the selection of the most efficient estimator is an empirical question that depends on the system under investigation.

In this study, we sought to identify the most statistically efficient estimator of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals we recorded the activity of nearly every neuron in volumes of cortex about 100 $\mu$m in diameter (150--350 cells).  We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with several shared latent factors to capture unobserved inputs and network fluctuations.  To test this hypothesis, we demonstrated that, in cross-validation tests, a covariance matrix estimator based on this correlation structure outperformed estimators regularized toward several other plausible low-dimensional approximations. We analyzed how the inferred network of interactions was related to the physical distances between cells and their orientation preferences.  Consistent with known synaptic connectivity, the density of positive interactions decreased rapidly with distance and with differences in preferred orientation, whereas negative interactions were less selective.  To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and local field potentials.

@@

The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(1),\ldots,x(n)$ as
\begin{equation}
C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
\end{equation}
where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent).  
The sample covariance matrix is unbiased, \emph{i.e.} $\E{C_{\sf 0}}=\Sigma$.
For finite sample sizes, however, $C_{\sf 0}$ is not as close to $\Sigma$ as a number of biased estimators.

@@

Finally, we consider the fourth estimator $C_{\sf sparse+lowrank}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D). This estimator has the form \cite{Chandrasekaran:2010,Ma:2013}
\begin{equation}
C_{\sf sparse+latent} = (S - LL^\T)^{-1}
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a $d\times p$ matrix of factor loadings. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.

@@

\section*{Discussion}
\subsection*{Big neural data}
\subsection*{Meaning of interactions}
\subsection*{Fitting probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
\subsection*{Future directions}

\section*{Methods}
% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\subsection*{Data Acquisition}
Fast 3D two-photon imaging of calcium signals \emph{in vivo} was performed as described in \cite{Cotton:2013}
\subsection*{Covariance Estimation}
TODO
\subsection*{Cross-validation}
TODO
\subsection*{Simulation}
TODO

\section*{Acknowledgments}
Genevera Allen
